### 2023-04-26 manuscript/narrative.rmd

---
title: "Narrative"
author: "Anton Olsson-Collentne"
date: '2023-03-14'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 6, fig.asp = 0.75)
```

In Olsson-Collentine et al., (2019) we demonstrate that variation in reliability between the studies included in a meta-analysis will lead to an overestimate of statistical heterogeneity. This point is also made by Hunter and Schmidt (year) whose claims are repeated in a recent paper published by Perspectives on Psychological Science (REF). However, in this article we demonstrate that this overestimate only holds in the absence of true heterogeneity (i.e., under a fixed effect meta-analytic framework). In the presence of true heterogeneity, imperfect reliability in the studies instead leads to an underestimate of the true heterogeneity.

## Method and results