### 2023-05-11 code/tables.qmd
table code file..

---
format: pdf
---

# Project: Unreliable heterogeneity

```{r setup, include=FALSE}
# Script purpose: create tables for manuscript
# code: Anton Olsson-Collentine

knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(kableExtra) #For creating good-looking pdf tables
```

## Table 1
### Variance in reliability leads to positive bias in heterogeneity estimates

```{r effect-size-heterogeneity-table}
#library(kableExtra)
rooted_relabilities <- matrix(rep(c(.6, .70,.80), 3),
                              ncol = 3, byrow = TRUE)
rho <- c(0.0, .30, .50)
ma <- c("I", "II", "III")

rooted_relabilities <- sweep(rooted_relabilities, 1, rho, "*")
devs <- apply(rooted_relabilities, 1, sd)
table1 <- data.frame(ma, rho, rooted_relabilities, devs)

positive_bias_table <- table1 %>%
         knitr::kable(booktabs = T, escape = FALSE, digits = 2,
          col.names = c("Meta-Analysis",
                        "$\\rho_{xy}$",
                        "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .6",
                        "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .7",
                        "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .8",
                        "SD(ES)")) %>%
  kable_styling(position = "left") %>%
  add_header_above(c(" " = 2, "Study 1"= 1, "Study 2" = 1, "Study 3" = 1)) %>%
  add_header_above(c(" " = 2, "Observed Effect Sizes" = 3)) %>%
  footnote(general = "Reproduced from Olsson Collentine et al. (2020). The values under Study 1, 2 and 3 are observed effect sizes for that study given its measurement reliability $\\\\sqrt{R_{xx'}}\\\\times\\\\sqrt{R_{yy'}}$ and the true effect size $\\\\rho_{xy}$ when within-study sample size is infinite. SD (ES) is the standard deviation of the observed effect sizes for meta-analysis I, II and III, equivalent to heterogeneity given infinite within-study sample sizes. Code to reproduce table: LINK",
  threeparttable = TRUE, escape = FALSE, footnote_as_chunk = TRUE)

positive_bias_table
```

## Table 2 (version 1)
### Imperfect reliability leads to negative bias in heterogeneity estimates

```{r negative-bias-heterogeneity}
library(kableExtra)

theta  <- matrix(rep(c(0.1, 0.15, 0.2), 3), ncol = 3, byrow = TRUE)
R <- c(1, 0.8, 0.6)
r <- sweep(theta, 1, R, "*")
sds <- apply(r, 1, sd)

dat <- cbind(as.character(R), r, sds)

dat %>% knitr::kable(booktabs = TRUE, escape = FALSE, digits = 2,
                    align=rep('c', 5),
 col.names = c("$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$",
               "Study 1",
               "Study 2",
               "Study 3",
               "SD(ES)")) %>%
 kable_styling(position = "left") %>%
 add_header_above(c(" " = 1, "Observed Effect Sizes" = 3)) %>%
 footnote(general = "The values under Study 1, 2 and 3 are observed effect sizes for that study given the measurement reliability $\\\\sqrt{R_{xx'}}\\\\times\\\\sqrt{R_{yy'}}$ and the true effect size of that study (first row) when within-study sample size is infinite. SD (ES) is the standard deviation of the observed effect sizes, equivalent to heterogeneity given infinite within-study sample sizes. Code to reproduce table: LINK",
  threeparttable = TRUE, escape = FALSE, footnote_as_chunk = TRUE)
```

## Table 2 (version 2)
### Imperfect reliability leads to negative bias in heterogeneity estimates

```{r, alternative negative bias table}
#Alternative table
rel <- c("$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = 1",
         "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = 0.8",
         "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = 0.6")

dat <- cbind(rel, r, sds)

negative_bias_table <- dat %>%
            knitr::kable(booktabs = TRUE, escape = FALSE, digits = 2,
            col.names = c("",
                        "Study 1",
                        "Study 2",
                        "Study 3",
                        "SD(ES)")) %>%
            kable_styling(position = "left") %>%
            add_header_above(c(" " = 1, "Observed Effect Sizes" = 3)) %>%
            column_spec (1, border_right = FALSE) %>%
            footnote(general = "The values under Study 1, 2 and 3 are observed effect sizes for that study given the measurement reliability $\\\\sqrt{R_{xx'}}\\\\times\\\\sqrt{R_{yy'}}$ and the true effect size of that study (first row) when within-study sample size is infinite. SD (ES) is the standard deviation of the observed effect sizes, equivalent to heterogeneity given infinite within-study sample sizes. Code to reproduce table: LINK",
  threeparttable = TRUE, escape = FALSE, footnote_as_chunk = TRUE)

negative_bias_table
```

```{r output tables}
#Saving the output of all tables so that they can be imported to the main manuscript easily

saveRDS(list(positve_bias_table = positive_bias_table,
             negative_bias_table = negative_bias_table),
        "../manuscript/tables.RDS")
```