### 2023-06-11 supplement/supplement_A.qmd
draft of supplement A + B

---
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Supplement A: varying reliability SD and the effect of N and k

```{r Figure A1, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_A_fig1.png")
```

```{r Figure A2, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_A_fig2.png")
```

```{r Figure A3, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_A_fig3.png")
```