### 2023-06-14 supplement/supplement_A.qmd
changed supplement names to match manuscript appearance order

---

format: pdf

---
# Supplement A: converting heterogeneity between Pearson's r and Fisher's z

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

The table below shows the results of our procedure to convert between studies variance expressed in Pearson's $r$ to between-studies variance expressed in Fisher's $z$ through the use of $I^2$ as described in the main text.

_Table A1_.
```{r}

source("../code/01_compute_tau_fishers_z.r") #all computations were made here

#Gives us the object 'cond'
cond_table <- cond[,c(1, 3, 2, 4, 6)]

knitr::kable(cond_table, booktabs = TRUE, col.names = c("$\\rho$", "N", "$\\tau_r$", "$I^2$", "$\\tau_z$"), escape = FALSE)
```

_Note._ $\rho$ = average effect size, N = (fixed) within-study sample size, $\tau_r$ = between-studies standard deviation in Pearson's $r$, $\tau_z$ between-studies standard deviation in Fisher's $z$. The $I^2$ statistic is here expressed as a proportion.