### 2023-06-14 supplement/supplement_B.qmd

---
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Supplement B: perfect reliability and Pearson's $r$

Figure 1 shows the bias in heterogeneity estimates in the absence of measurement unreliability in primary studies (i.e., reliability = 1).

```{r Figure B1, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_B_fig1.png")
```

_Figure B1._ Bias in heterogeneity estimates in the absence of unreliability in primary studies (i.e., reliability = 1). The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).

As can be seen, all three models show a positive bias when heterogeneity is zero, as heterogeneity estimates are truncated at zero. However, we also see a positive bias when heterogeneity is above zero for the Pearson's correlation using the HV method (second row). This is due to a dependancy between effect size and sampling variance for Pearson's $r$ as normally computed. That is, the sampling variance is estimated as $\hat{\sigma}^2_i = \frac{(1-r_i^2)^2}{n_i - 1}$. We do not see this effect for the HS method or HV method using Fisher's $z$. In the HS method $r_i$ has been replaced by $\bar{r}$ and for Fisher's $z$ the sampling variance is approximated as $\hat{\sigma}^2_i = 1 / (n_i -3)$, in both cases removing the dependency between effect size and sampling variance. To demonstrate that this is indeed the case, Figure 2 demonstrates how the bias in Pearson's $r$ disappears if we reshuffle the sampling variances among primary studies so that the dependency between effect size and sampling variance disappears.

```{r Figure B2, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_B_fig2.png")
```

_Figure B2._ Removal of the dependency between effect size and sampling variance removes the positive bias in heterogeneity estimates for larger true heterogeneity when using Pearson's $r$. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 1e3 repetitions. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).